Online {\IterComp}

### 2017-11-17 motivation.tex
minor

    Typical single-run performance metrics, like runtime or Instructions Per Cycle~(IPC), are not fit for purpose. To show this, we
    performed iterative compilation on \texttt{tiff2bw}, evaluating each optimization sequence on a different input and using either
    runtime or IPC as our efficiency metric. We selected a set of ten inputs from the KDataSets. When
    iterative compilation is driven by runtime, the selected optimization slows down the benchmark by 3\% compared to \texttt{-O3}. When
    driven by IPC, it is even worse with an average slowdown of 49\%. Its very high IPC of 3.25 does not indicate high performance. The
    actual best optimization over the whole input set improves performance by 7\% while only having an average IPC of 2.31.
   Higher IPC does not necessarily mean better performance,
 different optimizations affect IPC in different ways.
   Similarly, a shorter runtime may be due to inputs with fewer work to process.

\begin{table}[t]
\centering
\label{tab:motivation_metrics}
\begin{tabular}{|l|l|l|}
\hline
\textbf{Metric} & \textbf{Speedup/Slowdown} & \textbf{IPC} \\ \hline
\rowcolor{Gray} Oracle (Speedup)             & 7\% speedup                                 & 2.31                      \\ \hline
IPC                          & 49\% slowdown                                & 3.25                      \\ \hline
\rowcolor{Gray}1/Runtime                    & 3\% slowdown                                  & {-}   \\ \hline
\end{tabular}
\caption{Comparison of iterative compilation guided by different metrics.
The oracle is guided by the average speedup over \texttt{-O3} on all inputs.
For IPC and runtime, each optimization was evaluated on a different input.
}
\end{table}

>>>>

    Typical single-run performance metrics, like runtime or Instructions Per Cycle~(IPC), are not fit for purpose. To show this, we
    performed iterative compilation on \texttt{tiff2bw}, evaluating each optimization sequence on a different input and using either
    runtime or IPC as our efficiency metric. We selected a set of ten inputs from the KDataSets. When
    iterative compilation is driven by runtime, the selected optimization slows down the benchmark by 3\% compared to \texttt{-O3}. When
    driven by IPC, it is even worse with an average slowdown of 49\%. Its very high IPC of 3.25 does not indicate high performance. The
    actual best optimization over the whole input set improves performance by 7\% while only having an average IPC of 2.31.
    When the binary changes, higher IPC does not necessarily mean better performance, different optimizations affect IPC in different ways.
    Similarly, when the input changes, shorter runtimes do not mean faster execution, it may just be that processing the input requires less work.

\begin{table}[t]
\centering
\label{tab:motivation_metrics}
\begin{tabular}{|l|l|l|}
\hline
\textbf{Metric} & \textbf{Average Speedup/Slowdown} & \textbf{IPC} \\ \hline
\rowcolor{Gray} Oracle (Speedup)             & 7\% speedup                                 & 2.31                      \\ \hline
IPC                          & 49\% slowdown                                & 3.25                      \\ \hline
\rowcolor{Gray}1/Runtime                    & 3\% slowdown                                  & {-}   \\ \hline
\end{tabular}
\caption{Comparison of iterative compilation guided by different metrics.
The oracle is guided by the average speedup over \texttt{-O3} on all inputs.
For IPC and runtime, each optimization was evaluated on a different input.
}
\end{table}

### 2017-11-17 results.tex
minor formats
minor

We evaluate the quality of the selected optimization sequences by comparing their average speedup over \texttt{-O3} across all 1,000 inputs
of each benchmark.
% None of the configurations degrades the performance over \texttt{-O3} (with statistical significance). \FIXME{PP: According to the figure there are slowdowns!!!}
Figure~\ref{fig:speedups} shows these speedups for all test benchmarks. In most cases, \OraclePP is very close to \OracleRM, achieving on
average 80\% of its speedup. This demonstrates that it is a valid choice to drive iterative compilation decision using work efficiency. The
few cases where the two strategies produce noticeably different results are for the \texttt{tiff} benchmarks. The weaker correlation between
work and unoptimized runtime for these benchmarks is to blame for the lower speedups achieved by \OraclePP.

@@

\begin{figure*}[t]
  \centering
  \subfigure[\OptProf]{\includegraphics[width=0.3\textwidth]{figs/Opt-Profiling-deployment.pdf}}
  \subfigure[\WCRelax-5\%]{\includegraphics[width=0.3\textwidth]{figs/WC-Relax-5-deployment.pdf}}
  \subfigure[\WPRelax-5\%]{\includegraphics[width=0.3\textwidth]{figs/WP-Relax-5-deployment.pdf}}
  \caption{Evaluation of the profitability/cost experienced by the user throughout the search process (labeled as [Live]) and of the best optimization sequence selected by the search. These results provide evidence that the proposed approach is profitable in real-world online scenarios.}
  \label{fig:deployment_cost}
\end{figure*}

We also evaluate the actual cost perceived by the user when applying online iterative compilation.
To this end we used the architecture described in Section~\ref{sec:oic-infra} with the search based on genetic algorithm~\cite{knijnenburg02,cooper99,kulkarni04}.
Our results provide evidence that the proposed approach is profitable in real-world online scenarios.
Figure~\ref{fig:deployment_cost} shows, for each benchmark and configuration, the profitability of the final optimization sequence
selected by the search (as in Figure~\ref{fig:speedups}) but also the average profitability/cost experienced by the user throughout the search process (labeled as [Live]),
which takes into account the overhead of the work profiling.

With online iterative optimization, although the 
user may experience slowdown during the search process by executing sub-optimal versions of the program,
a Search Driver based on genetic algorithm is able to focus the search on promising optimization sequences, reducing the risk of
executing sub-optimal versions of the program as the search evolves through generations.
After the first generation, the optimization sequences produced tend to overcome this issue, with their benefits surpassing the profiling overheads.
We evaluate the deployment cost for the three configurations that use work profiling, namely \OptProf, \WCRelax-5\%, and \WPRelax-5\%.
In all three configurations, the online search has been shown profitable on average.
The wider search space is also able to improve the performance gains over \texttt{-O3}.

>>>>

We evaluate the quality of the selected optimization sequences by comparing their average speedup over \texttt{-O3} across all 1,000 inputs
of each benchmark.
Figure~\ref{fig:speedups} shows these speedups for all test benchmarks. In most cases, \OraclePP is very close to \OracleRM, achieving on
average 80\% of its speedup. This demonstrates that it is a valid choice to drive iterative compilation decision using work efficiency. The
few cases where the two strategies produce noticeably different results are for the \texttt{tiff} benchmarks. The weaker correlation between
work and unoptimized runtime for these benchmarks is to blame for the lower speedups achieved by \OraclePP.

@@

\begin{figure*}[t]
  \centering
  \subfigure[\OptProf]{\includegraphics[width=0.3\textwidth]{figs/Opt-Profiling-deployment.pdf}}
  \hfill
  \subfigure[\WCRelax-5\%]{\includegraphics[width=0.3\textwidth]{figs/WC-Relax-5-deployment.pdf}}
  \hfill
  \subfigure[\WPRelax-5\%]{\includegraphics[width=0.3\textwidth]{figs/WP-Relax-5-deployment.pdf}}
  \caption{Speedup over \texttt{-O3} for the best optimization selected by the search, as well as the average speedup as experienced by the
    user throughout the search process ([Live]). The results show that the proposed approach is profitable in real-world online scenarios.}
  \label{fig:deployment_cost}
\end{figure*}

A typical argument against online adaptive techniques is that the overhead of management and the slowdown due to testing suboptimal
configuration might overshadow the benefits of the technique. In this subsection, we establish that online iterative compilation is
profitable even when factoring in the online costs. We modify the architecture described in Section~\ref{sec:oic-infra} to make it more
realistic by using genetic search~\cite{knijnenburg02,kulkarni04} over the whole space of possible optimizations.

Figure~\ref{fig:deployment_cost} shows, for each benchmark and configuration, the speedup over \texttt{-O3} of the best optimization sequence
evaluated by the search but also the average profitability/cost experienced by the user throughout the search process (labeled as [Live]),
which takes into account the overhead of the work profiling.
With online iterative optimization, although the 
user may experience slowdown during the search process by executing sub-optimal versions of the program,
a Search Driver based on genetic algorithm is able to focus the search on promising optimization sequences, reducing the risk of
executing sub-optimal versions of the program as the search evolves through generations.
After the first generation, the optimization sequences produced tend to overcome this issue, with their benefits surpassing the profiling overheads.
We evaluate the deployment cost for the three configurations that use work profiling, namely \OptProf, \WCRelax-5\%, and \WPRelax-5\%.
In all three configurations, the online search has been shown profitable on average.
The wider search space is also able to improve the performance gains over \texttt{-O3}.
